---
title: "P8105_hw2_my2644"
author: "ymysherry"
date: "9/29/2020"
output: github_document
---

This is my solution to HW2.

```{r setup}
library(tidyverse)
library(readxl)
```

##Problem 1
#Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df =
  read_xlsx(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017
```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation",
  skip = 1, 
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018)  %>%
  relocate(year)
```
```{r}
precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation",
  skip = 1, 
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017)  %>%
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )


precip_df = 
bind_rows(precip_2018, precip_2017) 
left_join(precip_df, month_df, by = "month")
  
```

```{r}
 nrow(trashwheel_df)
```


This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trash collects that trash, and stores it in a dumpster. The dataset contains information on year, months, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.

Look for total precipitation in 2018 and median number of sports balls in a dumpster in 2017
```{r}
sum(pull(precip_2018,total),na.rm=TRUE)
```
The total precipiration in 2018 is 70.33 inches.

```{r}
sports_balls_2017_df= filter(trashwheel_df, year==2017)
median(pull(sports_balls_2017_df, sports_balls), na.rm= TRUE)
```
The median number of sports balls in a dumpster in 2017 is 8.


##Problem 2
#Read and clean the NYC subway dataset. Retain useful variables and convert the entry variable from character (YES vs NO) to a logical variable

```{r}
nycsubway_df =
  read_csv(
  "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")

nycsubway_df = janitor::clean_names(nycsubway_df) 

nycsubway_df = 
  select(nycsubway_df, line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada)

nycsubway_df =  
  mutate(nycsubway_df, entry = recode (entry, "YES"="TRUE","NO"="FALSE"), entry = as.logical(entry))
skimr::skim(nycsubway_df)
```
#Describe the dataset

This dataset contains information from the NYC subway transit data. The dataset contains 19 variables in total, including information on entrance, exit, station names, routes, lines, as well as specific longitude and latitude of subway stations. There are a total of `r nrow(nycsubway_df)` rows in our final dataset. Additional data cleaning has been performed by the janitor step. Entry variable has been converted from character to logical by using mutate. This is a tidy dataset with dimensions of `r nrow(nycsubway_df)*ncol(nycsubway_df)`. Number of columns represents variables and number of rows represents observations.

##Answering questions using the NYC subway transit data
#how many distinct stations (by name and line) are there
```{r}
distinct_stations_df = 
  distinct(nycsubway_df, line, station_name, .keep_all = TRUE)
```
1. There are `r nrow(distinct_stations_df)` distinct stations

#how many stations are ADA compliant
```{r}
ADAcompliant_df=
  filter(distinct_stations_df, ada == TRUE)
```
2. `r nrow(ADAcompliant_df)` stations are ADA compliant

#What proportion of station entrances / exits without vending allow entrance
```{r}
entrance_novending_df = 
  filter(nycsubway_df, vending == "NO", entry == TRUE) 

novending_df = 
  filter(nycsubway_df, vending == "NO")
```
3. The proportion of stationentrances/exits without vending allows entrance is `r nrow(entrance_novending_df)/nrow(novending_df)`

